---
title: "P8105 Homework 1"
author: "Rezwana Habib"
date: "2026-09-26"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

# Problem 1
```{r}

data("penguins", package = "palmerpenguins")

```

The penguins dataset contains `r nrow(penguins)` observations and `r ncol(penguins)` variables. It includes information on three penguin species (Adelie, Chinstrap, and Gentoo) from three islands (Biscoe, Dream, and Torgersen).

The dataset includes measurements such as bill length, bill depth, flipper length, and body mass, along with each penguin's species, island, sex, and year of observation.

The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.

```{r penguin_scatterplot}
ggplot(penguins, aes(
  x = bill_length_mm,
  y = flipper_length_mm,
  color = species
)) +
  geom_point()
ggsave("penguin_scatterplot.png", height = 4, width = 6)
```



# Problem 2


```{r create_dataframe}
set.seed(1234)

sample_df = tibble(
  numeric_var = rnorm(10),
  logical_var = numeric_var > 0,
  character_var = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J"),
  factor_var = factor(c("low", "medium", "high", "low", "medium",
                        "high", "low", "medium", "high", "low"))
)

sample_df
```
```{r calculate_means}
mean(pull(sample_df, numeric_var))
mean(pull(sample_df, logical_var))
mean(pull(sample_df, character_var))
mean(pull(sample_df, factor_var))
```

## Mean Results
The mean works for the numeric and logical variables. The logical variable has a mean of 0.4 because four out of ten values are TRUE, which R treats as 1. The character and factor variables return NA because they are not numeric or logical values.

```{r numeric_conversion, results='hide', warning=FALSE}
as.numeric(pull(sample_df, logical_var))

as.numeric(pull(sample_df, character_var))

as.numeric(pull(sample_df, factor_var))
```

## Numeric Conversion
The logical variable converts to 1s and 0s because R treats TRUE as 1 and FALSE as 0. This explains why its mean can be calculated. 

The character variable returns NA because letters cannot be converted into numbers. 

The factor variable converts to numeric codes representing its levels rather than the actual category names. 

These codes are just labels for categories, so calculating their mean would not be meaningful. 